Reading Notes

### 2020-07-24 outline.tex
comparing votca and grits with cognitive load lens

>>>>

<compare VOTCA's backmapping to my proposed scheme>
VOTCA makes some progress towards a TRUE simulation pipeline: they include a docker container with their pre-built software stack and example input files <see https://github.com/votca/votca -- using this code via docker> (but some of these input files don't work and the codecoverage on their repo as of time of writing is 0\%), and while including example inputs is a good first step, very little walkthrough (aside from code comments buried in the xml) is given. A manual is provided but there is some disconnect between the example files and those given in the manual <https://github.com/votca/csg-manual/releases>. the code is a mix of C++, shell commands, python, and interacts with compiled programs through the command-line and mpi. which makes their working a little tenuous. And judging by the amount of confusion in their mailing list <https://groups.google.com/forum/#!forum/votca> there is room for improvement in the API. Overall though this is program has many features, interfaces with many popular MD engines (lammps, gromacs, hoomd), and is under active, collaborative development.
A good question might be if I'm so into OSS and collaboration--why don't I just contribute an MSIBI package to VOTCA. This is a reasonable question. The answer is because I find VOTCA to be near-overwhelming in it's scope and the cognitive load of trying to fit this whole project into my brain is not ideal. I also don't like their API. by choosing a pure python API I will be able to make use of jupyter notebooks which can be run on cloud servers using tools like mybinder <link mybinder, link to tutorials I;ve made> which can combine executable code with markdown text and images. In this way a tutorial can provide exact commands needed to reproduce the example intimately connected with the rationale and justification of that example. I would like to think that my code speaks for itself, but after reading other people's code, I have a renewed appreciation for explanations in paragraph form, diagrams, and figures.

For an example of what I mean by cognitive load, let's compare an example of my proposed coarse-graining scheme to VOTCAs. In the following example we will take a hexane molecule and map it to three coarse grain sites (two carbon atoms per site). VOTCA uses an xml file to do this:
<insert hexane.xml>
It may be easy to think that because what we are doing--mapping two carbons and their attached hydrogens to a single bead so that six alkyl carbons becomes A-B-A--that the process will also be simple. In order to specify a bead-mapping, we need to know the residue number, the residue name, and the atom name. For a six atom system this is not too bad, but this could get tedious and error-prone for larger systems. VOTCA's scheme also requires that you specify the bonding and the angles resulting from those bonds--I will discuss this more after discussing my scheme below:
\begin{lstlisting}
hexane_sites = [("_A", "CC[H3]"), ("_B", "CC")]
\end{lstlisting}
The result of using this code in the `coarse` function <link to grits> is a three-site hexane model A-B-A. The mapping here is specified as a list of bead name / smarts string pairs. Although they do add some element of cognitive load, SMARTS strings are designed to be human and machine readable and can match atoms based on their chemical environment. No need to look up the name or number of each residue simply say I want two alkyl carbons (CC) specifying that it is an alkyl chain tail by requiring one carbon to have three attached hydrogens (C[H3]). The bonding of the coarse grain beads is then inferred based on the bonding of the mapped atoms.
